---
title: "Practical Machine Learning"
author: "Yading Song"
date: "21 December 2014"
output: html_document
---

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Dataset
### Training
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

### Testing 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Task
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

## Loading data 

### Set path and load the correct working directory and necessary packages. 
```{r}
setwd("~/Google Drive/JobApplication/DataScicentist/courses/08_PracticalMachineLearning/writeUp")
library(caret)
```

### Load the data for training and testing 
We set the seed to 1234 for reproduciblity. 

```{r}
# "classe" is the target variable  
set.seed(1234)
data <- read.csv("pml-training.csv")
```

### Clean the data 
Since there is no enough points (NA and empty cells) in the dataset. We removed these variables in the preprocessing procedure. 

```{r}
factor_names <- names(data)
observations <- nrow(data)
noise <- vector()
count <- 1
# remove NA data points 
for (i in seq(1:160)){
    na_number <- sum(is.na(data[factor_names[i]]))    
    if (na_number > observations/2 ){
        noise[count] <- factor_names[i]
        count <- count + 1 
        #print(factor_names[i])  
    }
}
```

### remove NULL cells 
```{r}
redundent <- c("raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
null_part1 <- c("kurtosis_roll_belt", "kurtosis_picth_belt", "kurtosis_yaw_belt", 
                "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt", 
                "max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt")
null_part2 <- c("kurtosis_roll_arm", "kurtosis_picth_arm", "kurtosis_yaw_arm", 
                "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm")
null_part3 <- c("kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", 
                "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell", "skewness_pitch_dumbbell", 
                "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell", "amplitude_yaw_dumbbell")
null_part4 <- c("kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm", 
                "skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", 
                "max_yaw_forearm", "min_yaw_forearm", "amplitude_yaw_forearm")
noises <- c(noise, redundent, null_part1, null_part2, null_part3, null_part4)
```

### Obtain filtered data for analysis 
As the first couple columns are not related with prediction. (time stamps, names, etc.) We will remove these from further analysis.

```{r}
my_vars <- factor_names %in% noises 
clean_data <- data[!my_vars]
clean_data <- clean_data[, 3:55]
```

### Explore the data
boxplot showing the difference in different levels.
```{r}
boxplot(accel_belt_y~classe, data=clean_data, xlab="Levels A-E", ylab="Accel_belt_y")
```

Correlation analyses are carried out. 
```{r}
correlations <- abs(cor(clean_data[,-53]))
diag(correlations) <- 0
which(correlations > 0.8,arr.ind=T)
```

### PCA to reduce dimensionality 
25 components are needed to retain 95% of the variance
```{r}
preProc <- preProcess(clean_data[, -53], method="pca", thresh=0.95)
print(preProc)

### Separate training and testing data
```{r}
# cross validation 70% for training 
inTrain <- createDataPartition(y=clean_data$classe, p=0.7, list=FALSE)
training <- clean_data[inTrain, ]
testing <- clean_data[-inTrain, ]
```

## Building the machine learning model 
```{r}
# Method 1: We are using decision tree 
modFit1 <- train(classe~., method="rpart", preProcess="pca", data=training)
print(modFit1)
# Method 2: We are using lda
modFit2 <- train(classe~., method="lda", preProcess="pca", data=training)
print(modFit2)
```

## Examine the results 
```{r}
result1 <- predict(modFit1, newdata=testing)
print(table(result1, testing$classe))
# In sample error 
sum(result1==testing$classe)/nrow(testing)
result2 <- predict(modFit2, newdata=testing)
print(table(result2, testing$classe))
sum(result2==testing$classe)/nrow(testing)
```

## In Sample and Out of Sample (OOS) Error 
The results show that the prediction accuracy is about 75%. Out of sample error can be drawn from a different data with other participants (e.g., age). Also, this dataset was collected from people who did little exercise. It can bring noises if experienced athletes are doing the same experiment. In addition, we used only 53 variables out of 120-ish because of lack of data. If the measurement is complete, it will be more precise. 

## Testing other data 
```{r}
validation <- read.csv("pml-testing.csv")
validate1 <- predict(modFit1, newdata=validation)
validate2 <- predict(modFit2, newdata=validation)
print("Using rpart")
print(validate1)
print("Using lda")
print(validate2)

```

## Report summary 
In this report, we use only methods "rpart" and "lda" with some data preprocessing. However, empty and some data points were removed due to the data. The error rate is about 25% (baseline 80%). Other methods should be applied in the future combining more machine learning techniques. 

## Examine the results 
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```